---
title: "HW4: Hodge Podge Em"
author: "Mike Keating"
format: pdf
editor: visual
---

## Task 1: Conceptual Questions

### Q1: What is the purpose of the lapply() function? What is the equivalent purrr function?

The purpose of lapply() is to apply a function to each element of a list.

The equivalent purr function is map().

### Q2: Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall") on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify method = "kendall" when calling lapply()

```{r}

df1 <- data.frame(col1 = c(1 ,3, 4), col2 =c(2, 5, 5))
df2 <- df1 * 2

my_list <- list(df1, df2)

lapply(my_list, cor, method = "kendall")


```

### Q3: What are 2 advantages of using purr functions instead of BaseR apply family?

The purr functions allow for the use of helper/anonymous functions which will help us write more compact code.

The purr functions (map) have options like map_dbl() and map_int() can help us ensure our output is of a specific type.

### Q4: What is a side-effect function?

A side-effect function is a function that performs an action outside of it's normal scope. In other words, something else happens in addition to returning a value.

### Q5: Why can you name a variable sd in a function and and not cause any issues with the sd function?

The variables created within the function are local/accessible only to the function. They are not initialized in the greater environment, so they do not conflict with the sd function.

## Task  2: Writing R Functions

### Q1: getRMSE

Write a basic function (call it getRMSE()) that takes in a vector of responses and a vector of predictions and outputs the RMSE.

```{r}
getRMSE <- function(resp, pred, ...){
  args <- list(...)
  # Show a warning if NA is present
  if (anyNA(resp)) {
    if ("na.rm" %in% names(args)){
      if (isFALSE(args$na.rm)){
        warning("NA found in response vector. Consider setting na.rm = TRUE")
        }
      } else {
        warning("NA found in response vector. Consider setting na.rm = TRUE")
        }  
  }
  
  # Get squared errors
  squared_errors <- (resp - pred)^2
  
  # Get mean squared errors, mse
  mse <- mean(squared_errors, ...)
  
  # Finally, get rmse
  rmse <- sqrt(mse)
  
  return(rmse)
  
  
}
```

### Q2: Test getRMSE

Create response values and predictions

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

```

Test RMSE function

```{r}
getRMSE(resp, pred)

```

Repeat after replacing two of the response values with missing values.

```{r}
# To make things simple let's just replace the first two values in the response values with NA
resp_NA <- replace(resp, 1:2, NA_real_)


# Calling without remove na arument
getRMSE(resp_NA, pred)
```

```{r}
# Calling with our optional argument
getRMSE(resp_NA, pred, na.rm = TRUE)
```

### Q3: getMAE()

```{r}
getMAE <- function(resp, pred, ...){
   args <- list(...)
  # Show a warning if NA is present
  if (anyNA(resp)) {
    if ("na.rm" %in% names(args)){
      if (isFALSE(args$na.rm)){
        warning("NA found in response vector. Consider setting na.rm = TRUE")
        }
      } else {
        warning("NA found in response vector. Consider setting na.rm = TRUE")
        }  
  }
  # Get absolute error
   ae <- abs(resp - pred)
   mae <- mean(ae, ...)
   
   return(mae)
  
}
```

### Q4:  Test getMAE()

Create response values and predictions

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test MAE function:

```{r}
getMAE(resp,pred)
```

Repeat after replacing two of the response values with missing values.

```{r}
# To make things simple let's just replace the first two values in the response values with NA
resp_NA <- replace(resp, 1:2, NA_real_)


# Calling without remove na arument
getMAE(resp_NA, pred)
```

```{r}
getMAE(resp_NA, pred, na.rm = TRUE)
```

### Q5: Wrapper for Both Metrics

```{r}
getMetrics <- function(resp, pred, metrics = c("rmse", "mae"), ...){
  if (!is.numeric(resp) | !is.numeric(pred)){
    stop("ERROR: Responses or predictions not a numeric vector")
  }
  
  results = list()
  
  
  
  if ("rmse" %in% metrics){
    results$rmse <- getRMSE(resp, pred, ...)
    
  }
  if ("mae" %in% metrics){
    results$mae <- getMAE(resp, pred, ...)
  }
  
  if (length(results) == 0){
    stop("ERROR: Unknown metrics supplied")
  }
  
  return(results)
  
}
```

### Q6: Test New Metric Function

Call function for RMSE:

```{r}
getMetrics(resp, pred, c("rmse"))
```

Call function for MAE:

```{r}
getMetrics(resp, pred, c("mae"))
```

Calling function for both metrics:

```{r}
getMetrics(resp, pred, c("rmse","mae"))
```

Repeat with NA values:

```{r}
# Recall we already replaced with NA in our resp_NA object
getMetrics(resp_NA, pred, c("rmse", "mae"))
```

```{r}
# Recall we already replaced with NA in our resp_NA object
getMetrics(resp_NA, pred, c("rmse", "mae"), na.rm = TRUE)
```
